\chapter{Strategy and Testing} \label{cap:testing}

The testing strategy for this web application is organized around the current test suite structure: unit tests for service behavior, REST integration tests for the Spring MVC and security layer, provider integration tests for third-party event discovery, and end-to-end tests for complete user journeys. This gives the project coverage at the most important behavioral boundaries while keeping the tests deterministic and independent from live external services.

\section{Testing Approach}

The suite follows a layered approach:
\begin{itemize}
    \item \textbf{Unit tests} isolate service classes with Mockito. These tests check business rules directly, such as meeting proposal validation, invite status transitions, user registration validation, discovery result merging, iCalendar rendering, and Spring Security user loading.
    \item \textbf{REST integration tests} start the Spring Boot application context with an H2 in-memory database and use \texttt{MockMvc}. These tests verify controller routes, security behavior, CSRF protection, redirects, model attributes, rendered views, and persistence side effects.
    \item \textbf{Third-party provider integration tests} replace each provider's HTTP client with a local JSON server. This allows the tests to verify URL query parameters and JSON mapping without depending on Ticketmaster, SeatGeek, or Agenda Cultural de Lisboa being available during the build.
    \item \textbf{End-to-end tests} are integrated in the suite to validate complete authentication and meeting workflows from the user's perspective.
\end{itemize}

The main tools used are JUnit 5, Mockito, Spring Boot Test, Spring Security Test, MockMvc, Selenium through HtmlUnitDriver, and H2. H2 is used by the REST integration tests and by the end-to-end tests so that repository operations and transaction behavior are exercised with a concrete database that is separate from the production database.

To support consistent testing, Continuous Integration (CI) was set up using GitHub Actions. This allows the same Maven test command to run automatically when changes are made to the codebase.

\section{Test Design Process}

The tests were created feature by feature. For each feature, the expected behavior was tested to confirm that the intended flow was present. After that, tests were added for error situations that already existed in the application, such as invalid credentials, duplicate registration, invalid meeting dates, missing CSRF tokens, unknown invitees, and invalid response actions.

Once the main success and error paths were covered, the relevant input fields and service parameters for each feature were reviewed. When a field could receive malformed or missing data, tests for that case were added and the implementation was adjusted when the behavior was unsafe or unclear. This is how validation cases such as blank usernames, blank passwords, missing meeting titles, null dates, null invitee lists, and malformed discovered-event intervals were added to the suite.

\section{Assignment Requirements Mapping}

The implemented suite maps to the assignment requirements as follows:

\begin{itemize}
    \item \textbf{Mocks:} unit tests use Mockito to isolate services from repositories and providers.
    \item \textbf{Third-party integrations:} use local JSON servers to test provider HTTP requests and response mapping without relying on live external APIs.
    \item \textbf{REST API integration tests:} use \texttt{MockMvc} against the Spring MVC and security configuration.
    \item \textbf{Integration tests with a concrete database:} implemented by the REST and end-to-end support classes, which configure H2 in-memory databases with \texttt{create-drop} schema management instead of using the production database.
    \item \textbf{E2E tests:} use Selenium's \texttt{HtmlUnitDriver}, a random-port Spring Boot server, and an H2 test database.
    \item \textbf{Continuous Integration:} implemented with a GitHub Actions workflow that runs \texttt{mvn -B clean test}.
\end{itemize}

\section{Unit Test Coverage}

The unit tests are concentrated in \texttt{src/test/java/unit}. They focus on service logic and avoid loading the full Spring application context.

The current unit coverage includes:
\begin{itemize}
    \item \textbf{MeetingService}: proposing meetings, rejecting invalid titles or times, treating a null invitee list as empty, rejecting unknown invitees, accepting or rejecting invitations, retrieving calendar and pending invite data, and copying discovered events into the calendar.
    \item \textbf{UserService}: creating users, trimming usernames, rejecting blank usernames and passwords, rejecting duplicate usernames, and reporting unknown users.
    \item \textbf{DiscoveryService}: returning configured providers, ignoring blank searches, skipping unconfigured providers, deduplicating results, and sorting events by start time.
    \item \textbf{ICalService}: rendering calendar headers and events, escaping special characters, omitting blank descriptions, mapping participant statuses, and handling null email values.
    \item \textbf{AppUserDetailsService}: loading Spring Security user details and throwing the expected exception when a username does not exist.
\end{itemize}

\section{REST Integration Coverage}

The REST integration tests are concentrated in \texttt{src/test/java/integration/rest}. They load the Spring application with an H2 database and use \texttt{MockMvc} to exercise the web layer.

The current REST coverage includes:
\begin{itemize}
    \item registration and login flows, including duplicate usernames, invalid credentials, redirects, password hashing, and CSRF rejection;
    \item meeting proposal and response flows, including participant persistence, accept and decline actions, invalid response actions, and invalid date ranges;
    \item calendar rendering, including authenticated user data, meetings, pending invites, and generated iCalendar subscription links;
    \item discovery rendering and copying a discovered event into a user's calendar using a fake configured provider;
    \item iCalendar feed download, including successful \texttt{text/calendar} output for a valid token and a \texttt{404} response for an unknown token.
\end{itemize}

\section{Provider Integration Coverage}

The provider integration tests use a lightweight local HTTP server to return controlled JSON responses. This keeps the tests repeatable while still testing the real HTTP request construction and response mapping logic.

The current provider coverage includes:
\begin{itemize}
    \item Agenda Cultural de Lisboa response mapping, including occurrence dates, Lisbon timezone conversion, time parsing, venue extraction, and HTML stripping from descriptions;
    \item SeatGeek response mapping, including client id query parameters and UTC datetime parsing;
    \item Ticketmaster response mapping, including API key and country query parameters, nested \texttt{\_embedded} event data, venue extraction, and ISO instant parsing.
\end{itemize}

\section{End-to-End Coverage}

The suite also includes end-to-end coverage for the main user journeys. These tests complement the unit and integration tests by validating the behavior as complete flows rather than isolated classes or controller calls.

The current end-to-end coverage includes:
\begin{itemize}
    \item authentication flows, including registration and login behavior;
    \item meeting flows, including proposing meetings, displaying calendar state, and responding to invitations;
    \item persistence of user-facing changes across the service and repository layers during the tested workflows.
\end{itemize}

Together, these tests cover both successful and unsuccessful paths, such as duplicate registration, invalid credentials, missing CSRF tokens, invalid meeting dates, unknown response actions, blank service inputs, unknown invitees, and malformed discovered event intervals.
